Why aren't significance tests commonly used for linear regression diagnostics?

### 2022-09-13 paper_comparison/paper_comparison/paper_comparison.Rmd
update visual inference

Graphical summaries in which residuals are plotted against fitted values or other functions of the predictor variables that are approximately orthogonal to residuals are referred to as standard residual plots in @cook1982residuals. As shown in Figure [ref here], the top-left panel is a good residual plot with residuals evenly distributed at both sides of the horizontal zero line showing no noticeable patterns. There are various types of departures from a good residual plot. We will only discuss three most commonly checked departures in this paper. Namely, non-linearity, heterskedasticity and non-normality.

@@

Heterskedasticity refers to the presence of nonconst error variance in a regression model. It is mostly due to the strict but false assumptions on the variance-covariance matrix of the error term. The usual pattern of heterskedasticity on a residual plot is the inconsistent spread of the residuals at different x values. Visually, it sometimes results in the so-called "butterfly" shape as shown in the bottom-left panel of Figure [ref here], or the "left-triangle" and "right-triganle" shape where the smallest variance occurs at the edges of the x-axis.

Non-normality is usually harder to be detected from a residual plot compared to non-linearity and heterskedasticity. Besides, a quantile-quantile plot can often do a better job for this task. Note that not all regression models assume normality of the error term, but a certain amount of them does. If this assumption happens to be false, then it is expected to see data points do not distribute noramlly on the y-axis. For example, given a skewed error distribution, one will see fewer data points and more outliers on one side of the zeroline as shown in the bottom-right of Figure [ref here].

@@

Example residual plots given in \ref{fig:commondepartures} are examined by the corresponding F-test, Breusch-Pagan test and Shapiro--Wilk test as shown in \ref{tab:example-residual-plot-table}. In the example, both the Breusch-Pagan test and the Shapiro--Wilk test rejects the null hypothesis for departures they do not intend to examine. As discussed in @cook1982residuals, most residual-based tests for a particular type of departure from model assumptions are sensitive to other types of departures. It is likely the null hypothesis is correctly rejected but for the wrong reason, which is known as the "Type III error". Additionally, outliers will often incorrectly trigger the rejection of the null hypothesis despite the residuals are well-behaved [@cook_applied_1999]. This can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers.

@@

-   Very briefly explain what a lineup test is

-   Lineup (maybe using your previous residual plot for nonlinearity)

-   $p$-value calulcation, briefly, including how multiple selections are handled

-   power calculation, briefly, including how multiple selections are handled

Visual inference was first introduced in a 1999 Joint Statistical Meetings (JSM) talk with the title "Inference for Data Visualization" by @buja_inference_1999 as an idea to address the issue of valid inference for visual discoveries of data plots [@gelman_exploratory_2004]. Later, in the Bayesian context, data plots was systematically considered as model diagnostics by taking advantage of the data simulated from the assumed statistical models [@gelman_bayesian_2003; @gelman_exploratory_2004].

It was surprising that the essential components of visual inference had actually been established in @buja_inference_1999, but it was not until 10 years later that @buja_statistical_2009 formalized it as an inferential framework to extend confirmatory statistics to visual discoveries. This framework redefines the test statistics, tests, null distribution, significance levels and $p$-value for visual discovery modelled on the confirmatory statistical testing. Figure \ref{fig:parallelism} outlines the parallelism between conventional tests and visual discovery.

```{=html}
<!-- ![Parallelism between multiple quantitative testing and visual discovery [@buja_statistical_2009]. Visible features in a plot are viewed as a collection of test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$, and any visual discoveries that are inconsistent with the null hypothesis are treated as evidence against the null. For regression diagnostics, the null hypothesis would be the assumed model, and visual discoveries could be any visual features in favour of any alternatives. \label{fig:parallelism}](figures/rsta2009012001.jpg){width=450 height=341}
-->
```
In visual inference, a collection of test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$ is defined, where $\boldsymbol{\mathrm{y}}$ is the data and $I$ is a set of all possible visual features. @buja_statistical_2009 described each of the test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})$ as a measurement of the degree of presence of a visual feature. Alternatively, @majumder_validation_2013 avoids the use of visual features and defined the visual statistics $T(.)$ as a mapping from a dataset to a data plot. Both definitions of visual test statistics are valid, but in the rest of the paper the first definition will be used as it covers some details needed by the following discussion. A visual discovery is defined as a rejection of a null hypothesis, and the same null hypothesis can be rejected by many different visual discoveries [@buja_statistical_2009]. For regression diagnostics, the null hypothesis would be the assumed model, while the visual discoveries would be any findings that are inconsistent with the null hypothesis. The same regression model can be rejected by many reasons with residual plot, including non-linearity and heteroskedasticity as shown in Figure \ref{fig:residual-plot-cubic-heter}.

```{r residual-plot-cubic-heter, fig.cap='Residuals vs. fitted values plot for a classical linear regression model. The residuals are produced by fitting a two-predictor multiple linear regression model with data generated from a cubic linear model. From the residual plot, "butterfly shape" can be observed which generally would be interpretd as evidence of heteroskedasticity. Further, from the outline of the shape, nonlinear patterns exist. Both visual discoveries are evidence against the null hypothesis, though heteroskedasticity actually does not exist in the data generating process. \\label{fig:residual-plot-cubic-heter}'}

@@

### Sampling from the null distribution {#se:sampling-from-null}

The null distribution of plots refers to the infinite collection of plots of null datasets sampled from $H_0$. It is defined as the analogue of the null distribution of test statistics in conventional test [@buja_statistical_2009]. In practice, a finite number of plots of null datasets could be generated, called null plots. In the context of regression diagnostics, sampling data from $H_0$ is equivalent to sampling data from the assumed model. As @buja_statistical_2009 suggested, $H_0$ is usually composited by a collection of distributions controlled by nuisance parameters. Since regression models can have various forms, there is no general solution to this problem, but it sometimes can be reduced to so called "reference distribution" by applying one of the three methods: (i) sampling from a conditional distribution given a minimal sufficient statistic under $H_0$, (ii) parametric bootstrap sampling with nuisance parameters estimated under $H_0$, and (iii) Bayesian posterior predictive sampling.

The conditional distribution given a minimal sufficient statistic is the best justified reference distribution among the three [@buja_statistical_2009]. Suppose there exists a minimal sufficient statistic $\boldsymbol{S}(\boldsymbol{y})$ under the null hypothesis, any null datasets $\boldsymbol{y^{*}}$ should fulfil the condition $\boldsymbol{S}(\boldsymbol{y}) = \boldsymbol{s}$. Using the classical normal linear regression model as example, the minimal sufficient statistic is $\boldsymbol{S}(\boldsymbol{y}) = (\hat{\boldsymbol{\beta}}, \boldsymbol{e}'\boldsymbol{e})$, where $\hat{\boldsymbol{\beta}}$ are the coefficient estimators and $\boldsymbol{e}'\boldsymbol{e}$ is the residual sum of square. Alternatively, the minimal sufficient statistic can be constructed as $\boldsymbol{S}(\boldsymbol{y}) = (\hat{\boldsymbol{y}}, ||\boldsymbol{e}||)$, where $\hat{\boldsymbol{y}}$ are the fitted values and $||\boldsymbol{e}||$ is the length of residuals, which is more intuitive as suggested by @buja_statistical_2009. Since the fitted values are held fixed, the variation can only occur in the residual space. And because the length of residual is also held fixed, residuals obtained from a null dataset has to be a random rotation of $\boldsymbol{e}$ in the residual space. With this property, null residuals can be simulated by regressing $N$ i.i.d standard normal random draws on the regressors, then rescaling it by the ratio of residual sum of square in two regressions.

>>>>

Graphical summaries in which residuals are plotted against fitted values or other functions of the predictor variables that are approximately orthogonal to residuals are referred to as standard residual plots in @cook1982residuals. As shown in Figure \\ref{fig:commondepartures}, the top-left panel is a good residual plot with residuals evenly distributed at both sides of the horizontal zero line showing no noticeable patterns. There are various types of departures from a good residual plot. We will only discuss three most commonly checked departures in this paper. Namely, non-linearity, heterskedasticity and non-normality.

@@

Heterskedasticity refers to the presence of nonconst error variance in a regression model. It is mostly due to the strict but false assumptions on the variance-covariance matrix of the error term. The usual pattern of heterskedasticity on a residual plot is the inconsistent spread of the residuals at different x values. Visually, it sometimes results in the so-called "butterfly" shape as shown in the bottom-left panel of Figure \\ref{fig:commondepartures}, or the "left-triangle" and "right-triganle" shape where the smallest variance occurs at the edges of the x-axis.

Non-normality is usually harder to be detected from a residual plot compared to non-linearity and heterskedasticity. Besides, a quantile-quantile plot can often do a better job for this task. Note that not all regression models assume normality of the error term, but a certain amount of them does. If this assumption happens to be false, then it is expected to see data points do not distribute noramlly on the y-axis. For example, given a skewed error distribution, one will see fewer data points and more outliers on one side of the zeroline as shown in the bottom-right of Figure \ref{fig:commondepartures}.

@@

Example residual plots given in Figure \ref{fig:commondepartures} are examined by the corresponding F-test, Breusch-Pagan test and Shapiro--Wilk test as shown in Table \ref{tab:example-residual-plot-table}. In the example, both the Breusch-Pagan test and the Shapiro--Wilk test rejects the null hypothesis for departures they do not intend to examine. As discussed in @cook1982residuals, most residual-based tests for a particular type of departure from model assumptions are sensitive to other types of departures. It is likely the null hypothesis is correctly rejected but for the wrong reason, which is known as the "Type III error". Additionally, outliers will often incorrectly trigger the rejection of the null hypothesis despite the residuals are well-behaved [@cook_applied_1999]. This can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers.

@@

Visual inference was first introduced in a 1999 Joint Statistical Meetings (JSM) talk with the title "Inference for Data Visualization" by @buja_inference_1999 as an idea to address the issue of valid inference for visual discoveries of data plots [@gelman_exploratory_2004]. Later, @buja_statistical_2009 proposed the lineup protocol as a visual test inspired by the "police lineup" or "identity parade" which is the act of asking the eyewitness to identify criminal suspect from a group of irrelevant people. The protocol consists of $m$ randomly placed plots, where one plot is the actual data plot, and the remaining $m - 1$ plots have the identical graphical production as the data plot except the data has been replaced with data consistent with the null hypothesis. Then, an observer who have not seen the actual data plot will be asked to point out the most different plot from the lineup. Figure \ref{fig:first-example-lineup} is an example of a lineup protocol. If the actual data plot at position 11 is identifiable, then it is evidence for the rejection of the null hypothesis that the regression model is correctly specified. In fact, the example lineup is generated from a misspecified regression model with non-linearity issue. So one may notice the cubic curve in the actual residual plot.

Simulation of the null data used in the 19 null plots are

```{r first-example-lineup, fig.cap="Example lineup consists of 20 randomly placed residual plots, where 19 are null plots with data consistent with the null hypothesis, and one is the actual residual plot. Can you find the most different plot?"}
mod <- poly_model(include_z = TRUE, sigma = 1)
POLY_MODEL$plot_lineup(mod$gen_lineup(300, pos = 11), theme = theme_light(), remove_axis = TRUE, remove_grid_line = TRUE)
```

<!-- -   Very briefly explain what a lineup test is -->

<!-- -   Lineup (maybe using your previous residual plot for nonlinearity) -->

<!-- -   $p$-value calulcation, briefly, including how multiple selections are handled -->

<!-- -   power calculation, briefly, including how multiple selections are handled -->

<!-- Visual inference was first introduced in a 1999 Joint Statistical Meetings (JSM) talk with the title "Inference for Data Visualization" by @buja_inference_1999 as an idea to address the issue of valid inference for visual discoveries of data plots [@gelman_exploratory_2004]. Later, in the Bayesian context, data plots was systematically considered as model diagnostics by taking advantage of the data simulated from the assumed statistical models [@gelman_bayesian_2003; @gelman_exploratory_2004]. -->

<!-- It was surprising that the essential components of visual inference had actually been established in @buja_inference_1999, but it was not until 10 years later that @buja_statistical_2009 formalized it as an inferential framework to extend confirmatory statistics to visual discoveries. This framework redefines the test statistics, tests, null distribution, significance levels and $p$-value for visual discovery modelled on the confirmatory statistical testing. Figure \ref{fig:parallelism} outlines the parallelism between conventional tests and visual discovery. -->

```{=html}
<!-- ![Parallelism between multiple quantitative testing and visual discovery [@buja_statistical_2009]. Visible features in a plot are viewed as a collection of test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$, and any visual discoveries that are inconsistent with the null hypothesis are treated as evidence against the null. For regression diagnostics, the null hypothesis would be the assumed model, and visual discoveries could be any visual features in favour of any alternatives. \label{fig:parallelism}](figures/rsta2009012001.jpg){width=450 height=341}
-->
```

<!-- In visual inference, a collection of test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$ is defined, where $\boldsymbol{\mathrm{y}}$ is the data and $I$ is a set of all possible visual features. @buja_statistical_2009 described each of the test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})$ as a measurement of the degree of presence of a visual feature. Alternatively, @majumder_validation_2013 avoids the use of visual features and defined the visual statistics $T(.)$ as a mapping from a dataset to a data plot. Both definitions of visual test statistics are valid, but in the rest of the paper the first definition will be used as it covers some details needed by the following discussion. A visual discovery is defined as a rejection of a null hypothesis, and the same null hypothesis can be rejected by many different visual discoveries [@buja_statistical_2009]. For regression diagnostics, the null hypothesis would be the assumed model, while the visual discoveries would be any findings that are inconsistent with the null hypothesis. The same regression model can be rejected by many reasons with residual plot, including non-linearity and heteroskedasticity as shown in Figure \ref{fig:residual-plot-cubic-heter}. -->

```{r residual-plot-cubic-heter, fig.cap='Residuals vs. fitted values plot for a classical linear regression model. The residuals are produced by fitting a two-predictor multiple linear regression model with data generated from a cubic linear model. From the residual plot, "butterfly shape" can be observed which generally would be interpretd as evidence of heteroskedasticity. Further, from the outline of the shape, nonlinear patterns exist. Both visual discoveries are evidence against the null hypothesis, though heteroskedasticity actually does not exist in the data generating process. \\label{fig:residual-plot-cubic-heter}', eval = FALSE}

@@

<!-- ### Sampling from the null distribution {#se:sampling-from-null} -->

<!-- The null distribution of plots refers to the infinite collection of plots of null datasets sampled from $H_0$. It is defined as the analogue of the null distribution of test statistics in conventional test [@buja_statistical_2009]. In practice, a finite number of plots of null datasets could be generated, called null plots. In the context of regression diagnostics, sampling data from $H_0$ is equivalent to sampling data from the assumed model. As @buja_statistical_2009 suggested, $H_0$ is usually composited by a collection of distributions controlled by nuisance parameters. Since regression models can have various forms, there is no general solution to this problem, but it sometimes can be reduced to so called "reference distribution" by applying one of the three methods: (i) sampling from a conditional distribution given a minimal sufficient statistic under $H_0$, (ii) parametric bootstrap sampling with nuisance parameters estimated under $H_0$, and (iii) Bayesian posterior predictive sampling. -->

<!-- The conditional distribution given a minimal sufficient statistic is the best justified reference distribution among the three [@buja_statistical_2009]. Suppose there exists a minimal sufficient statistic $\boldsymbol{S}(\boldsymbol{y})$ under the null hypothesis, any null datasets $\boldsymbol{y^{*}}$ should fulfil the condition $\boldsymbol{S}(\boldsymbol{y}) = \boldsymbol{s}$. Using the classical normal linear regression model as example, the minimal sufficient statistic is $\boldsymbol{S}(\boldsymbol{y}) = (\hat{\boldsymbol{\beta}}, \boldsymbol{e}'\boldsymbol{e})$, where $\hat{\boldsymbol{\beta}}$ are the coefficient estimators and $\boldsymbol{e}'\boldsymbol{e}$ is the residual sum of square. Alternatively, the minimal sufficient statistic can be constructed as $\boldsymbol{S}(\boldsymbol{y}) = (\hat{\boldsymbol{y}}, ||\boldsymbol{e}||)$, where $\hat{\boldsymbol{y}}$ are the fitted values and $||\boldsymbol{e}||$ is the length of residuals, which is more intuitive as suggested by @buja_statistical_2009. Since the fitted values are held fixed, the variation can only occur in the residual space. And because the length of residual is also held fixed, residuals obtained from a null dataset has to be a random rotation of $\boldsymbol{e}$ in the residual space. With this property, null residuals can be simulated by regressing $N$ i.i.d standard normal random draws on the regressors, then rescaling it by the ratio of residual sum of square in two regressions. -->

### 2022-09-13 paper_comparison/paper_comparison/paper_comparison.tex
update visual inference

Graphical summaries in which residuals are plotted against fitted values
or other functions of the predictor variables that are approximately
orthogonal to residuals are referred to as standard residual plots in
\citet{cook1982residuals}. As shown in Figure {[}ref here{]}, the
top-left panel is a good residual plot with residuals evenly distributed
at both sides of the horizontal zero line showing no noticeable
patterns. There are various types of departures from a good residual
plot. We will only discuss three most commonly checked departures in
this paper. Namely, non-linearity, heterskedasticity and non-normality.

@@

Heterskedasticity refers to the presence of nonconst error variance in a
regression model. It is mostly due to the strict but false assumptions
on the variance-covariance matrix of the error term. The usual pattern
of heterskedasticity on a residual plot is the inconsistent spread of
the residuals at different x values. Visually, it sometimes results in
the so-called ``butterfly'' shape as shown in the bottom-left panel of
Figure {[}ref here{]}, or the ``left-triangle'' and ``right-triganle''
shape where the smallest variance occurs at the edges of the x-axis.

Non-normality is usually harder to be detected from a residual plot
compared to non-linearity and heterskedasticity. Besides, a
quantile-quantile plot can often do a better job for this task. Note
that not all regression models assume normality of the error term, but a
certain amount of them does. If this assumption happens to be false,
then it is expected to see data points do not distribute noramlly on the
y-axis. For example, given a skewed error distribution, one will see
fewer data points and more outliers on one side of the zeroline as shown
in the bottom-right of Figure {[}ref here{]}.

@@

Example residual plots given in \ref{fig:commondepartures} are examined
by the corresponding F-test, Breusch-Pagan test and Shapiro--Wilk test
as shown in \ref{tab:example-residual-plot-table}. In the example, both
the Breusch-Pagan test and the Shapiro--Wilk test rejects the null
hypothesis for departures they do not intend to examine. As discussed in
\citet{cook1982residuals}, most residual-based tests for a particular
type of departure from model assumptions are sensitive to other types of
departures. It is likely the null hypothesis is correctly rejected but
for the wrong reason, which is known as the ``Type III error''.
Additionally, outliers will often incorrectly trigger the rejection of
the null hypothesis despite the residuals are well-behaved
\citep{cook_applied_1999}. This can be largely avoided in diagnostic
plots as experienced analysts can evaluate the acceptability of
assumptions flexibly, even in the presence of outliers.

@@

Unlike hypothesis testing built upon rigorous statistical procedures,
reading diagnostic plots relies on graphical perception - human's
ability to interpret and decode the information embedded in the graph
{[}@cleveland\_graphical\_1984{]}, which is to some extent subjective
and indecisive. Further, visual discovery suffers from its unsecured and
unconfirmed nature where the degree of the presence of the visual
features typically can not be measured quantitatively and objectively,
which may lead to over or under-interpretations of the data. One such
example is finding an over-interpretation of the separation between gene
groups in a two-dimensional projection from a linear discriminant
analysis when in fact there are no differences in the expression levels
between the gene groups and separation is not an uncommon occurrence
{[}@roy\_chowdhury\_using\_2015{]}.

\begin{itemize}
\item
  Very briefly explain what a lineup test is
\item
  Lineup (maybe using your previous residual plot for nonlinearity)
\item
  \(p\)-value calulcation, briefly, including how multiple selections
  are handled
\item
  power calculation, briefly, including how multiple selections are
  handled
\end{itemize}

Visual inference was first introduced in a 1999 Joint Statistical
Meetings (JSM) talk with the title ``Inference for Data Visualization''
by \citet{buja_inference_1999} as an idea to address the issue of valid
inference for visual discoveries of data plots
\citep{gelman_exploratory_2004}. Later, in the Bayesian context, data
plots was systematically considered as model diagnostics by taking
advantage of the data simulated from the assumed statistical models
\citep{gelman_bayesian_2003, gelman_exploratory_2004}.

It was surprising that the essential components of visual inference had
actually been established in \citet{buja_inference_1999}, but it was not
until 10 years later that \citet{buja_statistical_2009} formalized it as
an inferential framework to extend confirmatory statistics to visual
discoveries. This framework redefines the test statistics, tests, null
distribution, significance levels and \(p\)-value for visual discovery
modelled on the confirmatory statistical testing. Figure
\ref{fig:parallelism} outlines the parallelism between conventional
tests and visual discovery.

In visual inference, a collection of test statistics
\(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\) is defined, where
\(\boldsymbol{\mathrm{y}}\) is the data and \(I\) is a set of all
possible visual features. \citet{buja_statistical_2009} described each
of the test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})\) as a
measurement of the degree of presence of a visual feature.
Alternatively, \citet{majumder_validation_2013} avoids the use of visual
features and defined the visual statistics \(T(.)\) as a mapping from a
dataset to a data plot. Both definitions of visual test statistics are
valid, but in the rest of the paper the first definition will be used as
it covers some details needed by the following discussion. A visual
discovery is defined as a rejection of a null hypothesis, and the same
null hypothesis can be rejected by many different visual discoveries
\citep{buja_statistical_2009}. For regression diagnostics, the null
hypothesis would be the assumed model, while the visual discoveries
would be any findings that are inconsistent with the null hypothesis.
The same regression model can be rejected by many reasons with residual
plot, including non-linearity and heteroskedasticity as shown in Figure
\ref{fig:residual-plot-cubic-heter}.

\begin{figure}
\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/residual-plot-cubic-heter-1} \caption{Residuals vs. fitted values plot for a classical linear regression model. The residuals are produced by fitting a two-predictor multiple linear regression model with data generated from a cubic linear model. From the residual plot, "butterfly shape" can be observed which generally would be interpretd as evidence of heteroskedasticity. Further, from the outline of the shape, nonlinear patterns exist. Both visual discoveries are evidence against the null hypothesis, though heteroskedasticity actually does not exist in the data generating process. \label{fig:residual-plot-cubic-heter}}\label{fig:residual-plot-cubic-heter}
\end{figure}

\hypertarget{se:sampling-from-null}{%
\subsubsection{Sampling from the null
distribution}\label{se:sampling-from-null}}

The null distribution of plots refers to the infinite collection of
plots of null datasets sampled from \(H_0\). It is defined as the
analogue of the null distribution of test statistics in conventional
test \citep{buja_statistical_2009}. In practice, a finite number of
plots of null datasets could be generated, called null plots. In the
context of regression diagnostics, sampling data from \(H_0\) is
equivalent to sampling data from the assumed model. As
\citet{buja_statistical_2009} suggested, \(H_0\) is usually composited
by a collection of distributions controlled by nuisance parameters.
Since regression models can have various forms, there is no general
solution to this problem, but it sometimes can be reduced to so called
``reference distribution'' by applying one of the three methods: (i)
sampling from a conditional distribution given a minimal sufficient
statistic under \(H_0\), (ii) parametric bootstrap sampling with
nuisance parameters estimated under \(H_0\), and (iii) Bayesian
posterior predictive sampling.

The conditional distribution given a minimal sufficient statistic is the
best justified reference distribution among the three
\citep{buja_statistical_2009}. Suppose there exists a minimal sufficient
statistic \(\boldsymbol{S}(\boldsymbol{y})\) under the null hypothesis,
any null datasets \(\boldsymbol{y^{*}}\) should fulfil the condition
\(\boldsymbol{S}(\boldsymbol{y}) = \boldsymbol{s}\). Using the classical
normal linear regression model as example, the minimal sufficient
statistic is
\(\boldsymbol{S}(\boldsymbol{y}) = (\hat{\boldsymbol{\beta}}, \boldsymbol{e}'\boldsymbol{e})\),
where \(\hat{\boldsymbol{\beta}}\) are the coefficient estimators and
\(\boldsymbol{e}'\boldsymbol{e}\) is the residual sum of square.
Alternatively, the minimal sufficient statistic can be constructed as
\(\boldsymbol{S}(\boldsymbol{y}) = (\hat{\boldsymbol{y}}, ||\boldsymbol{e}||)\),
where \(\hat{\boldsymbol{y}}\) are the fitted values and
\(||\boldsymbol{e}||\) is the length of residuals, which is more
intuitive as suggested by \citet{buja_statistical_2009}. Since the
fitted values are held fixed, the variation can only occur in the
residual space. And because the length of residual is also held fixed,
residuals obtained from a null dataset has to be a random rotation of
\(\boldsymbol{e}\) in the residual space. With this property, null
residuals can be simulated by regressing \(N\) i.i.d standard normal
random draws on the regressors, then rescaling it by the ratio of
residual sum of square in two regressions.

>>>>

Graphical summaries in which residuals are plotted against fitted values
or other functions of the predictor variables that are approximately
orthogonal to residuals are referred to as standard residual plots in
\citet{cook1982residuals}. As shown in Figure
\textbackslash ref\{fig:commondepartures\}, the top-left panel is a good
residual plot with residuals evenly distributed at both sides of the
horizontal zero line showing no noticeable patterns. There are various
types of departures from a good residual plot. We will only discuss
three most commonly checked departures in this paper. Namely,
non-linearity, heterskedasticity and non-normality.

@@

Heterskedasticity refers to the presence of nonconst error variance in a
regression model. It is mostly due to the strict but false assumptions
on the variance-covariance matrix of the error term. The usual pattern
of heterskedasticity on a residual plot is the inconsistent spread of
the residuals at different x values. Visually, it sometimes results in
the so-called ``butterfly'' shape as shown in the bottom-left panel of
Figure \textbackslash ref\{fig:commondepartures\}, or the
``left-triangle'' and ``right-triganle'' shape where the smallest
variance occurs at the edges of the x-axis.

Non-normality is usually harder to be detected from a residual plot
compared to non-linearity and heterskedasticity. Besides, a
quantile-quantile plot can often do a better job for this task. Note
that not all regression models assume normality of the error term, but a
certain amount of them does. If this assumption happens to be false,
then it is expected to see data points do not distribute noramlly on the
y-axis. For example, given a skewed error distribution, one will see
fewer data points and more outliers on one side of the zeroline as shown
in the bottom-right of Figure
\textbackslash ref\{fig:commondepartures\}.

@@

Example residual plots given in Figure \ref{fig:commondepartures} are
examined by the corresponding F-test, Breusch-Pagan test and
Shapiro--Wilk test as shown in Table
\ref{tab:example-residual-plot-table}. In the example, both the
Breusch-Pagan test and the Shapiro--Wilk test rejects the null
hypothesis for departures they do not intend to examine. As discussed in
\citet{cook1982residuals}, most residual-based tests for a particular
type of departure from model assumptions are sensitive to other types of
departures. It is likely the null hypothesis is correctly rejected but
for the wrong reason, which is known as the ``Type III error''.
Additionally, outliers will often incorrectly trigger the rejection of
the null hypothesis despite the residuals are well-behaved
\citep{cook_applied_1999}. This can be largely avoided in diagnostic
plots as experienced analysts can evaluate the acceptability of
assumptions flexibly, even in the presence of outliers.

@@

Unlike hypothesis testing built upon rigorous statistical procedures,
reading diagnostic plots relies on graphical perception - human's
ability to interpret and decode the information embedded in the graph
\citep{cleveland_graphical_1984}, which is to some extent subjective and
indecisive. Further, visual discovery suffers from its unsecured and
unconfirmed nature where the degree of the presence of the visual
features typically can not be measured quantitatively and objectively,
which may lead to over or under-interpretations of the data. One such
example is finding an over-interpretation of the separation between gene
groups in a two-dimensional projection from a linear discriminant
analysis when in fact there are no differences in the expression levels
between the gene groups and separation is not an uncommon occurrence
\citep{roy_chowdhury_using_2015}.

Visual inference was first introduced in a 1999 Joint Statistical
Meetings (JSM) talk with the title ``Inference for Data Visualization''
by \citet{buja_inference_1999} as an idea to address the issue of valid
inference for visual discoveries of data plots
\citep{gelman_exploratory_2004}. Later, \citet{buja_statistical_2009}
proposed the lineup protocol as a visual test inspired by the ``police
lineup'' or ``identity parade'' which is the act of asking the
eyewitness to identify criminal suspect from a group of irrelevant
people. The protocol consists of \(m\) randomly placed plots, where one
plot is the actual data plot, and the remaining \(m - 1\) plots have the
identical graphical production as the data plot except the data has been
replaced with data consistent with the null hypothesis. Then, an
observer who have not seen the actual data plot will be asked to point
out the most different plot from the lineup. Figure
\ref{fig:first-example-lineup} is an example of a lineup protocol. If
the actual data plot at position 11 is identifiable, then it is evidence
for the rejection of the null hypothesis that the regression model is
correctly specified. In fact, the example lineup is generated from a
misspecified regression model with non-linearity issue. So one may
notice the cubic curve in the actual residual plot.

Simulation of the null data used in the 19 null plots are

\begin{figure}
\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/first-example-lineup-1} \caption{Example lineup consists of 20 randomly placed residual plots, where 19 are null plots with data consistent with the null hypothesis, and one is the actual residual plot. Can you find the most different plot?}\label{fig:first-example-lineup}
\end{figure}
